---
title: "Introduction to AustralianPoliticians"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{australianpoliticians}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Default usage

Load the package with:

```{r setup}
library(AustralianPoliticians)
```

The purpose of `AustralianPoliticians` is to ...

## Show data request codes

Request codes are used to access the specific datasets in `AustralianPoliticians`. These codes used in conjunction with the `get_auspol()` function allow for a dataset to be downloaded and assigned a variable name. The `show_datacodes()` function prints these codes to the console as a tibble.

```{r}
show_datacodes()
```


## Download Australian Politician data

Each of the shown codes is associated with a .csv dataset. The `get_auspol()` function is used to download these datasets, allowing them to be assigned to a variable. To function correctly, the argument passed to `get_auspol()` must be a character string. If `get_auspol()` is called without being assigned to a variable it will print out a preview to the requested dataset to the console, as seen below.

```{r, eval = F}
get_auspol("all")
# A tibble: 1,781 x 20                                                                    
   uniqueID  surname allOtherNames firstName commonName
   <chr>     <chr>   <chr>         <chr>     <chr>     
 1 Abbott18~ Abbott  Richard Hart~ Richard   NA        
 2 Abbott18~ Abbott  Percy Phipps  Percy     NA        
 3 Abbott18~ Abbott  Macartney     Macartney Mac       
 4 Abbott18~ Abbott  Charles Lydi~ Charles   Aubrey    
 5 Abbott18~ Abbott  Joseph Palmer Joseph    NA        
 6 Abbott19~ Abbott  Anthony John  Anthony   Tony      
 7 Abel1939  Abel    John Arthur   John      NA        
 8 Abetz1958 Abetz   Eric          Eric      NA        
 9 Adams1943 Adams   Judith Anne   Judith    NA        
10 Adams1951 Adams   Dick Godfrey~ Dick      NA        
# ... with 1,771 more rows, and 15 more variables:
#   displayName <chr>, earlierOrLaterNames <chr>,
#   title <chr>, gender <chr>, birthDate <date>,
#   birthYear <dbl>, birthPlace <chr>,
#   deathDate <date>, member <dbl>, senator <dbl>,
#   wasPrimeMinister <dbl>, wikidataID <chr>,
#   wikipedia <chr>, adb <chr>, comments <chr>

all_auspol <- get_auspol("all")
```

## House of Representatives and Senate

In some cases it may be necessary to join datasets to gain more information than is contained in just one dataset. The `get_reps_senate()` function does this in downloading the `house of representatives` data or `senate` data and joins these with the `all` dataset to show the dates served by a politician in either political sector. It can do this for both datasets, joining `house of representatives` and `senate` to `all`, or for just one of `house of representative` or `senate` depending on what is required.

This function also includes an argument that allows for the associated request codes to be printed to the console.

```{r eval = F}
# Return codes used to call datasets
get_reps_senate("codes")
# A tibble: 3 x 2
  request_code dataset                                 
  <chr>        <chr>                                   
1 reps_senate  Generates a dataset of HoRs and Senate ~
2 reps         Generates a dataset of HoRs members     
3 senate       Generates a dataset of Senate members 

# Request HoR and Senate dataset
reps_senate <- get_reps_senate("reps_senate")

# Preview dataset
head(reps_senate)
# A tibble: 6 x 4
  uniqueID   from       to         house
  <chr>      <date>     <date>     <chr>
1 Abbott1869 1913-05-31 1919-11-03 HoR  
2 Abbott1886 1925-11-14 1929-10-12 HoR  
3 Abbott1886 1931-12-19 1937-03-28 HoR  
4 Abbott1891 1940-09-21 1949-10-31 HoR  
5 Abbott1957 1994-03-26 2019-05-18 HoR  
6 Abel1939   1975-12-13 1977-11-10 HoR


# Request HoR dataset
reps <- get_reps_senate("reps")

# Preview dataset
head(reps)
# A tibble: 6 x 4
  uniqueID   from       to         house
  <chr>      <date>     <date>     <chr>
1 Abbott1869 1913-05-31 1919-11-03 HoR  
2 Abbott1886 1925-11-14 1929-10-12 HoR  
3 Abbott1886 1931-12-19 1937-03-28 HoR  
4 Abbott1891 1940-09-21 1949-10-31 HoR  
5 Abbott1957 1994-03-26 2019-05-18 HoR  
6 Abel1939   1975-12-13 1977-11-10 HoR


# Request Senate dataset
senate <- get_reps_senate("senate")

#Preview dataset
head(senate)
# A tibble: 6 x 4
  uniqueID    from       to         house 
  <chr>       <date>     <date>     <chr> 
1 Abbott1859  1928-12-18 1929-06-30 Senate
2 Abbott1869  1925-11-14 1929-06-30 Senate
3 Abbott1877  1935-07-01 1941-06-30 Senate
4 Abetz1958   1994-02-22 NA         Senate
5 Adams1943   2005-07-01 2012-03-31 Senate
6 Adamson1857 1920-07-01 1922-05-02 Senate
```
